---
title: 'Exercise 2: R Markdown for Gapminder Exploration'
author: "Carleena Ortega"
date: "September 27 2019"
output:
  pdf_document:
    toc: yes
  html_document:
    keep_md: yes
    theme: yeti
    toc: yes
always_allow_html: yes
---

```{r, warning=FALSE} 
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
```

# Data on Gapminder

In this document, we will explore basic R functions on the gapminder data.

## Population of Countries
Here we show that gapminder has information on life expectancy, population, and GDP per capita of 142 countries from 1952
```{r}
str(gapminder)
```

## Summary of Gapminder Data

We summarize the gapminder data and show that life expectancy, populations and GDP per capita were collected between 1952 and 2007 in 5 continents. 
```{r}
summary(gapminder)
```

## Number of Countries per continent in 1952
Here we show the number of countries per continent that data was collected from when the study started in 1952.
```{r}
gapminder %>% 
  arrange(country)%>%
  filter(year == 1952) %>% 
  mutate(continent = fct_infreq(continent)) %>% 
  ggplot(aes(continent)) +
  geom_bar(aes(colour=4,fill=4)) +
  theme_light()
```


