---
title: 'HW2: Explore Gapminder and use dplyr'
author: "Carleena Ortega"
date: "27/09/2019"
output:
  pdf_document:
  html_document:
    keep_md: yes
    theme: cerulean
    toc: yes
  word_document:
    toc: yes
  always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
```

# Exercise 1
## 1.1 Filter
Use `filter()` to subset the `gapminder` data to three countries of your choice in the 1970’s.
```{r}
filtered <- gapminder %>%
  arrange(year) %>%
  filter(year > 1969, year < 1980, country == "Canada" | country == "Mexico" 
    | country == "Brazil") %>%
  arrange(country)
knitr::kable(filtered)
```


## 1.2 Pipe Operator
Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

```{r}
filtered %>%
  select(country, gdpPercap) 
```

## 1.3 Drop in Life Expectancy
Filter gapminder to all entries that have experienced a drop in life expectancy. Be sure to include a new variable that's the increase in life expectancy in your tibble. Hint: you might find the lag() or diff() functions useful.

```{r}
gapminder %>%
  group_by(country) %>%
  arrange(country,year) %>%
  mutate(change_LE=lifeExp-lag(lifeExp)) %>%
  filter(change_LE<0) 

```

## 1.4 Max

Choose one of the following:

**Filter gapminder so that it shows the max GDP per capita experienced by each country. 
    Hint: you might find the max() function useful here.**

OR

Filter gapminder to contain six rows: the rows with the three largest GDP per capita, 
 and the rows with the three smallest GDP per capita. Be sure to not create any 
intermediate objects when doing this (with, for example, the assignment operator). 
Hint: you might find the sort() function useful, or perhaps even the dplyr::slice() function.


```{r}
gapminder %>%
  group_by(country) %>%
  arrange(country,gdpPercap) %>%
  filter(gdpPercap==max(gdpPercap))
```

## 1.5 Scatterplot

Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. Ensure GDP per capita is on a log scale.

```{r}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(aes(x=log(gdpPercap),lifeExp)) +
  scale_x_log10(limits=c(9,11)) +
  geom_point(colour="blue") +
  labs(x="log(GDP per capita)",
    y="Life Expectancy", 
    title="Canada's Life Expectancy Increases with GDP per capita") +
  theme_bw()
```

# Exercise 2

Pick one categorical variable and one quantitative variable to explore. Answer the following questions in whichever way you think is appropriate, using dplyr:
What are possible values (or range, whichever is appropriate) of each variable?
What values are typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.
Feel free to use summary stats, tables, figures.

## 2.1 Gapminder Continent

For this exercise, we will use `continent` as a categorical variable and `pop` as quantitative variable from the  `gapminder` data set.

We can see that there are 5 continents: **Africa, Asia, Europe, Americas, and Oceania**
 Africa has the most number of continents and Oceania with the fewest countries.
```{r}
  gapminder %>% 
  arrange(country)%>%
  filter(year == 1952) %>% 
  mutate(continent = fct_infreq(continent)) %>% 
  ggplot(aes(continent)) +
  geom_bar(colour="3",fill="3") +
  labs(x="Continent",
    y="Number of Countries", 
    title="Number of Countries per Continent") +
  theme_bw()
```

We can also view this data as a table:
```{r}
gapminder %>% 
  filter(year == 1952) %>% 
  group_by(continent) %>% 
  summarize(number_of_countries = mean(length(country)))
```

## 2.2 Gapminder Pop

The mean population of the continents over the years are shown below:
```{r}
gapminder %>% 
  group_by(continent) %>%
  summarize(mean_popuplation=mean(pop))
```

The standard error of the population per country is shown below:
```{r}
gapminder %>%
  group_by(continent) %>%
  summarize(stdE_pop = sd(pop)/sqrt(n()))
```

The ranges of population for each continent is shown below:
```{r}
gapminder %>%
  group_by(continent) %>%
  summarize(min(pop),max(pop))
```

This shows that Asia has the most population at 1,318,683,096 and Africa had the least at 60,011. Hence, the range of population amongst all continents is from 60,011 to 1,318,683,096

# Exercise 3
  
  Make two plots that have some value to them. That is, plots that someone might actually consider making for an analysis. Just don’t make the same plots we made in class – feel free to use a data set from the datasets R package if you wish.

    A scatterplot of two quantitative variables.
    One other plot besides a scatterplot.

You don’t have to use all the data in every plot! It’s fine to filter down to one country or a small handful of countries.

## 3.1 Scatterplot

For this exercise, we will explore the relationship between the population and the life expectancy of Canada.

```{r}
gapminder %>%
  filter(country=="Canada") %>%
  ggplot(aes(pop, lifeExp, pop)) +
  geom_point()+
<<<<<<< HEAD
  labs(y="Life Expectancy", x="Population", title="The Life Expectancy of 
    Canadians with respect to Population") +
=======
  labs(y="Life Expectancy", x="Population", title="The Life Expectancy of 
    Canadians with respect to Population") +
>>>>>>> 1ca241f8288cec53a6f6659df35e11a33da8fe2c
  theme_light()

```

From this plot, we can observe the increase in Canadian life expectancy as population increases. This may be due to the increase in caretakers available for the aged population, more people pursue research that promote longevity and improved quality of life.

## 3.2 Smooth plot
```{r, warning = FALSE}
gapminder %>%
  filter(country=="Canada") %>%
  ggplot(aes(lifeExp, pop)) +
  geom_smooth()+
<<<<<<< HEAD
  labs(x="Life Expectancy", y="Population", title="The Life Expectancy of 
    Canadians with respect to Population") +
=======
  labs(x="Life Expectancy", y="Population", title="The Life Expectancy of 
    Canadians with respect to Population") +
>>>>>>> 1ca241f8288cec53a6f6659df35e11a33da8fe2c
  theme_light()

```

# Bonus

## Bonus 1

For people who want to take things further.

Evaluate this code and describe the result. Presumably the analyst’s intent was to get the data for Rwanda and Afghanistan. Did they succeed? Why or why not? If not, what is the correct way to do this?

filter(gapminder, country == c("Rwanda", "Afghanistan"))

```{r}
x<-filter(gapminder, country == c("Rwanda", "Afghanistan")) 
knitr::kable(x)
```
Upon entering that code, the gapminder data for Rwanda and Afghanistan appear and the analyst can continue to work on it (e.g. via piping)

## Bonus 2

Present numerical tables in a more attractive form using knitr::kable() for small tibbles (say, up to 10 rows), and DT::datatable() for larger tibbles.

*Please refer to Exercise 1.1 and Bonus 1*