---
title: 'HW2: Explore Gapminder and use dplyr'
author: "Carleena Ortega"
date: "27/09/2019"
output:
  pdf_document: null
  html_document:
    keep_md: yes
    theme: cerulean
    toc: yes
  word_document:
    toc: yes
  always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(gapminder))
```

# Exercise 1
## 1.1 Filter
Use `filter()` to subset the `gapminder` data to three countries of your choice in the 1970’s.
```{r}

filtered <- gapminder %>%
  arrange(year) %>%
  filter(year > 1969, year < 1980, country == "Canada" | country == "Mexico" 
    | country == "Brazil") %>%
  arrange(country)
```


## 1.2 Pipe Operator
Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

```{r}
filtered %>%
  select(country, gdpPercap) 
```


  



#1.3
#Filter gapminder to all entries that have experienced a drop in life expectancy. Be sure to include a new variable that's the increase in life expectancy in your tibble. Hint: you might find the lag() or diff() functions useful