---
title: 'HW05: Factor and Figure Management'
author: "Carleena Ortega"
date: "14/10/2019"
output: 
  html_document:
    toc: true
  pdf_document:
    toc: true
---

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(scales))
```



# Exercise 1: Explain the value of the here::here package

In your own words, summarize the value of the here::here package in 250 words or fewer.

# Exercise 2: Factor management

We choose the `gapminder` dataset and select `continent` as the factor to explore. The code below determines whether gapminder's continent is a factor or not; how many levels it has and what these levels are. In addition, we also use the str() command to show a summary of the dataset, including how many observations (rows) we have.
```{r}
class(gapminder$continent)
nlevels(gapminder$continent)
levels(gapminder$continent)
str(gapminder)
```
We now know that `continent` is a factor with 5 levels named Africa, Americas, Asia, Europe, and Oceania. We also have 1704 rows or observations as shown by the str() function.

Our next step is to filter and drop levels. We filter out Oceania and then drop the unused factor levels in continent. 

```{r}
drop_Ocn<-gapminder %>% 
  filter(continent !="Oceania") %>% 
  droplevels()

drop_Ocn$continent %>% 
  nlevels() 

drop_Ocn$continent %>% 
  levels()

str(drop_Ocn)
```

We have successfully filtered out the data from Oceania as we now have only 4 levels in continent named Africa, Americas, Asia, and Europe. The str() function also shows that we have 1680 observations, which is less than the original dataset (1704 observations).

Our next goal is to explore the effects of re-leveling a factor with the modified gapminder data set. For this, we will use the summarized max gross domestic product (GDP) per capita of each continent.

```{r}
drop_Ocn %>% 
  group_by(continent) %>% 
  mutate("max GDP"=max(gdpPercap)) %>% 
  arrange(gdpPercap) %>% 
  ggplot()+
  geom_bar(aes(continent),colour="light blue", fill="light blue")+
  labs(x="Continent",y="Max GDP ($)",title="The Max GDP of continents (Original)")+
  theme_bw()
 
```

The arrange() command applies just to the tibble and did not affect the plot so the continents are arranged in the default alphabetic


```{r}
drop_Ocn %>% 
  select(continent,gdpPercap) %>% 
  group_by(continent) %>%
  summarize(maxGDP=max(gdpPercap)) %>% 
  ggplot(aes(x= fct_reorder(continent,maxGDP), y =maxGDP)) +
  geom_bar(stat = 'identity',colour="royal blue",fill="royal blue")+
  labs(x="Continent",y="Max GDP ($)",title="The Max GDP of continents (Reordered)")+
  scale_y_continuous(labels=comma)+
  theme_bw()
```

This is a more intuitive way of presenting the max GDP of each continent compared to the first plot. It shows the plots in an ascending order that makes it easy to see that Africa has the lowest GDP and Asia has the highest GDP. It also makes it easier to compare between continents.

# Exercise 3: File input/output (I/O)

For this exercise, we will modify the `gapminder` data set to show only data from Africa in 2007. We will then export this file using write_csv().

```{r}
gm_africa2007 <- gapminder %>% 
  filter(year == 2007, continent=="Africa")

write_csv(gm_africa2007,here::here("HW05","gm_africa2007.csv"))
```

We will now reload the exported file back into R using read_csv() and play around its factor levels.

```{r"}
read_csv(here::here("HW05","gm_africa2007.csv")) %>% 
  arrange(lifeExp) %>%
  ggplot()+
  geom_point(aes(x=lifeExp,y=country))+
  labs(x="Country",y="Life Expectancy (years)",title="The Life Expectancy of African countries in 2007")+
  theme_bw()
```
#height gg plot 
The original plot is arranged alphabetically by default. This is preferred for some tasks but for our purpose, we will rearrange the countries in an ascending order according to their life expectancies using fct_order().
```{r}
read_csv(here::here("./HW05/gm_africa2007.csv")) %>%
  ggplot()+
  geom_point(aes(x=lifeExp,y=fct_reorder(country,lifeExp)))+
  labs(x="Country",y="Life Expectancy (years)",title="The Life Expectancy of African countries in 2007")+
  theme_bw()+
  ggsave("African 2007 LE.png")
```
#remove slashes but use ""

The life expectancy of Africans in 2007 ranges from around 40 to 79 with Swaziland having the lowest life expectancy and Reunion with the highest life expectancy.

We have successfully exported, imported, and modified the gm_africa2007 file. Therefore, I can say that it has successfully passed being written to file and being read back in.

# Exercise 4: Visualization design

Task: Create a side-by-side plot and juxtapose your first attempt (show the original figure as-is) with a revised attempt after some time spent working on it and implementing principles of effective plotting principles. Comment and reflect on the differences.

I have chosen to revise the plot I made for HW02 Exercise 3.2. The code and plot are as follows:
```{r warning=FALSE}
gapminder %>%
  filter(country=="Canada") %>%
  ggplot(aes(lifeExp, pop)) +
  geom_smooth()+
  labs(x="Life Expectancy", y="Population", title="The Life Expectancy of 
    Canadians with respect to Population") +
  theme_light()
```

I noticed that the data points are not shown, the axes do not start at zero, and the line has both blue and grey colours which are not explained in a legend, and that geom_smooth is not the best functio to plot the data.

I will further improve the plot by applying the principles I have learned about effective plotting. The plot is as follows:
```{r warning=FALSE}
Can_LEpop<-gapminder %>%
  filter(country=="Canada")

Can_LEpop %>% 
  ggplot(aes(lifeExp, pop)) +
  geom_point(size=3)+
  geom_line()+
  xlim(72,80)+
  labs(x="Life Expectancy", y="Population", title="The Life Expectancy of 
    Canadians with respect to Population") +
  theme_light()
```


# Exercise 5: Writing figures to file (10%)

Task: Use ggsave() to explicitly save a plot to file. Include the exported plot as part of your repository and assignment.

Then, use ![Alt text] (/path/to/img.png) to load and embed that file into your report. You can play around with various options, such as:

Arguments of ggsave(), such as width, height, resolution or text scaling.
Various graphics devices, e.g. a vector vs. raster format.
Explicit provision of the plot object p via ggsave(..., plot = p). Show a situation in which this actually matters.

```{r}
Can_LEpop
  ggsave(filename = "LE of canadians wrt Pop", plot="Can_LEpop",device=png(),path="./HW05/")
```

