---
title: 'HW05: Factor and Figure Management'
author: "Carleena Ortega"
date: "14/10/2019"
output: 
  html_document:
    toc: true
  pdf_document:
    toc: true
---

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
```



# Exercise 1: Explain the value of the here::here package

In your own words, summarize the value of the here::here package in 250 words or fewer.

# Exercise 2: Factor management

We choose the `gapminder` dataset and select `continent` as the factor to explore. The code below determines whether gapminder's continent is a factor or not; how many levels it has and what these levels are.
```{r}
class(gapminder$continent)
nlevels(gapminder$continent)
levels(gapminder$continent)

str(gapminder)
```
We now know that `continent` is a factor with 5 levels named Africa, Americas, Asia, Europe, and Oceania. We also have 1704 rows or observations as shown by the str() function.

Our next step is to filter and drop levels. We filter out Oceania and then drop the unused factor levels in continent. 

```{r}
drop_Ocn<-gapminder %>% 
  filter(continent !="Oceania") %>% 
  droplevels()

drop_Ocn$continent %>% 
  nlevels() 

drop_Ocn$continent %>% 
  levels()

str(drop_Ocn)
```

We have successfully filtered out the data from Oceania as we now have only 4 levels in continent named Africa, Americas, Asia, and Europe. The str() function also shows that we have 1680 observations, which is less than the original dataset (1704 observations).

Our next goal is to explore the effects of re-leveling a factor with the gapminder data set.

```{r}
drop_Ocn %>% 
  group_by(continent) %>% 
  mutate("max GDP"=max(gdpPercap)) %>% 
  ggplot()+
  geom_bar(aes(continent),colour="light blue", fill="light blue")+
  labs(x="Continent",y=",ax GDP ($)",title="The Max GDP of continents (Original)")+
  theme_bw()
  
```

By default, the continents are arranged alphabetically. We can use the forcats package (fct_infreq  and fct_rev() commands) to arrange the continents by frequence and in an ascending order of max (gross domestic product) GDP, respectively.

```{r}
drop_Ocn %>% 
  group_by(continent) %>% 
  mutate("max GDP"=max(gdpPercap)) %>% 
  ggplot()+
  geom_bar(aes(fct_rev(fct_infreq(continent))),colour="royal blue", fill="royal blue")+
  labs(x="Continent",y="Max GDP ($)",title="The Max GDP of continents (Reordered)")+
  theme_bw()
  
```

This is a more appealing and understandable way of presenting the max GDP of each continent compared to the first plot.
Explore the effects of re-leveling a factor in a tibble by:

# Exercise 3: File input/output (I/O)
Task: Experiment with at least one of:

write_csv()/read_csv() (and/or TSV friends),
saveRDS()/readRDS(),
dput()/dget().
You are expected to create something new, probably by filtering or grouped-summarization of your dataset (for e.g., Singer, Gapminder, or another dataset), export it to disk and then reload it back in using one of the packages above. You should use here::here() for reading in and writing out.

With the imported data, play around with factor levels and use factors to order your data with one of your factors (i.e. non-alphabetically). For the I/O method(s) you chose, comment on whether or not your newly created file survived the round trip of writing to file then reading back in.

# Exercise 4: Visualization design (20%)
Rubric:

5%: Writing
5%: Accuracy
10%: Vis quality
Go back through your previous assignments and class participation activities and find figures you created prior to the last week of the course. Recreate at least one figure in light of something you learned in the recent class meetings about visualization design and color.

Task: Create a side-by-side plot and juxtapose your first attempt (show the original figure as-is) with a revised attempt after some time spent working on it and implementing principles of effective plotting principles. Comment and reflect on the differences.

# Exercise 5: Writing figures to file (10%)
5%: Accuracy
5%: Code quality
Task: Use ggsave() to explicitly save a plot to file. Include the exported plot as part of your repository and assignment.

Then, use ![Alt text] (/path/to/img.png) to load and embed that file into your report. You can play around with various options, such as:

Arguments of ggsave(), such as width, height, resolution or text scaling.
Various graphics devices, e.g. a vector vs. raster format.
Explicit provision of the plot object p via ggsave(..., plot = p). Show a situation in which this actually matters.
